Automatic Hadith Graphs using Arabic Morphological Analysis

### 2011-06-16 trunk/tex/tp.tex
corrected a bug resulting in not detecting آب
also added some words to lexicon of time.

%A subproblem of NLP is morphology.
Morphological analyzers %~\cite{Sughaiyer:04}
consider the internal structure of a word and compose it from
several {\em morphemes}. 
A morpheme is a {\em stem} or an {\em affix}.
An affix is a {\em prefix, suffix,} or an {\em infix}.
Prefixes and suffixes are attached to the beginning and end of the word respectively; 
while infixes introduce changes in the stem itself. 
Morphological analyzers have been extensively used as building blocks 
in many text mining applications~\cite{Sou07}.  In Arabic, morphological analyzers such as  
Buckwalter~\cite{Buckwalter:02},
Beesley~\cite{Beesley:01},
SAMA ~\cite{Kulick:10},
ElixirFM~\cite{Otakar:07}, 
AMIRA~\cite{Diab:07,Benajiba:07},
MAGEAD~\cite{Habash:05}, 
and MADA+TOKAN~\cite{Habash:09}
exist.
They preprocess text to remove diacritics, short vowels, such as 
\RL{ha}, \RL{hu}, \RL{hi}, \RL{TaN}, \RL{huN}, \RL{hiN}.~\footnote{In this document, 
we use the default ArabTeX transliteration style ZDMG.} Diacritics help to disambiguate the
meaning of a word. For example, \noVocRL{wrd} may mean ``rose"(\RL{ward}), ``came"(\fullvocalize\RL{warada}) 
or ``and he replied"(\RL{waradda}). We think that using diacritics in the analysis can highly increase accuracy.
Current analyzers take as input whitespace-delimited tokens. However, in Arabic word boundaries can be more complex.
Arabic letters can have up to 
four different forms
corresponding to their position in a word, i.e. beginning,
middle, end of word and separate forms. 
This allows the phrase
\noTrnoVocRL{il_A\nospace almdrsT}  with no delimiter within
to be visually recognizable
as two separate words \noVocRL{il_A} (to) and \noVocRL{almdrsT} (the school).
This happens because the first word \noVocRL{il_A} ends with
\noVocRL{_A} a non-connecting letter. We propose to solve this problem, commonly known as the run-on word problem, using online tokenization.
This also solves the issue of multi-word expressions such as \noVocRL{`abid al-la.tiyf} being analyzed as one unit -- here a compound proper name.

@@

%Similar to %the 
%MADA+TOKAN, Sarf uses
%the several valid morphological analyses as 
%%a richness that should 
%richness at a higher abstract level, and does not view them as 
%inaccuracies to be corrected. 
Sarf differs from MADA+TOKAN in that it provides an interface for the 
application to interfere in judging the analyses much earlier and
at a finer granularity level. 
%In Sarf, there is no need for a separate tokenizer such as
Sarf does not need a separate tokenizer such as
TOKAN since each solution keeps a stack of positions
that partition text into morphemes.
%The application-specific controller can build the trace 
%when needed.

@@

The main tasks of the thesis are listed below: \\
\begin{enumerate}
\item \textbf{Build and Enhance Sarf} \label{I:Sarf}
	\begin{enumerate}
	\item Port Buckwalter into C++ and mysql \label{I:portBuckwalter}
	\item Design and implement the Non-deterministic Finite State Automata framework \label{I:ndfsa}
	\item Implement an application-specific controller interface \label{I:applicationController}
	\item Support for disambiguation using diacritics \label{I:diacritics}
	\item Support for multi-word expressions \label{I:multiword}
	\item Solve the issue of `run-on' words \label{I:runon}
	\item Validate Sarf \label{I:validateSarf}
	\item Benchmark against existing corpora \label{I:benchMark}
	\end{enumerate}
\item \textbf{Support Sarf}
	\begin{enumerate}
	\item Optimize performance \label{I:optimizeSarf}
	\item Support for running Sarf under Windows \label{I:Sarfo}
	\item Refactor the code
	\item Write a user Manual
	\item Write a developer manual and documentation
	\end{enumerate}
\item \textbf{Build a time entity extraction application}
	\begin{enumerate}
	\item Recognize time entities
	\item Normalize extracted time entities
	\item Validate with a sorting application
	\end{enumerate}
\item \textbf{Automate Hadith Authentication}
	\begin{enumerate}
	\item Extend the Lexicon of Buckwalter with proper names 
	\item Implement the hadith segmentation tool
	\item Write the Narrator Equality Function
	\item Build Parial Order Graphs
	\item Correct Graph mistakes that result from Ambiguous narrator names which introduce false cycles to the graph
	\item Segment biography books using a graph coloring algorithm
	\item Extract Annotations from the biography books
	\item Assign annotations to graph nodes
	\item Enable and perform consistency checks between narrator chains
	\item Perform experiments on the effects of different narrators in hadith literature
	\item Write a user manual
	\item Distribute the tool to volunteer scholars and report their findings
	\end{enumerate}
\item \textbf{Write thesis Report and defend the thesis}
\end{enumerate}

@@

\begin{table}[bt]
\centering
\caption{Thesis Time Table}
\resizebox{1\columnwidth}{!}{
\begin{tabular}{|c|c||c|c||c|c||c||c|c|c|c||c|c|c|c||c|} \bottomrule[0.1em]
\textbf{Year} & \multicolumn{9}{|c|}{\textbf{2011}} & \multicolumn{6}{|c|}{\textbf{2012}} \\ \hline
\textbf{Semester} & ~ & \multicolumn{2}{|c||}{\textbf{\tiny Spring}} & \multicolumn{2}{|c||}{\textbf{\tiny Summer}} 
		& ~ & \multicolumn{4}{|c||}{\textbf{\tiny Fall}} & \multicolumn{4}{|c||}{\textbf{ \tiny Spring}} & ~ \\ \hline
\backslashbox{Task}{Month}
			& \textbf{Done} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8} & \textbf{9} & \textbf{\small 10} & 
			 \textbf{\small 11} & \textbf{\small 12} &
			 \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} &
				\begin{tabular}{c}
				\textbf{\small Future} \\
				\textbf{\small Work}
				\end{tabular}
			 \\ \toprule[0.1em] \bottomrule[0.1em]
\textbf{1.1)} & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{1.2)} & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ &\\ \hline
\textbf{1.3)} & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{1.4)} & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~  & ~ &\\ \hline
\textbf{1.5)} & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{1.6)} & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~  & ~ &\\ \hline
\textbf{1.7)} & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{1.8)} & \x & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ &\\ \hline
\textbf{1.9)} & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X & \X & ~ & \\ \hline
\textbf{1.10)}& ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X  & ~ & \\ \hline
\textbf{1.11)}& ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X & ~ & \\ \toprule[0.1em] \bottomrule[0.1em]
\textbf{2)}   & ~ & ~ & \X & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \toprule[0.1em] \bottomrule[0.1em]
\textbf{3.1)} & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{3.2)} & \x & ~ & ~ & ~ & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{3.3)} & \x & ~ & ~ & ~ & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{3.4)} & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{3.5)} & ~ & ~ & ~ & \X & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{3.6)} & ~ & ~ & ~ & \X & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{3.7)} & ~ & ~ & ~ & ~ & ~ & \X & \X & \X & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{3.8)} & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{3.9)} & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X & \X & \X & ~ & ~ & ~ & \X \\ \hline
\textbf{3.10)} & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X & \X & ~ & ~ & \X \\ \hline
\textbf{3.11)} & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X & ~ & \\ \hline
\textbf{3.12)} & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X \\ \toprule[0.1em] \bottomrule[0.1em]
\textbf{4)} & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X &  \\ \hline
\end{tabular}
}
\normalsize
\label{t:timetable}
\end{table}

>>>>

%A subproblem of NLP is morphology.
Morphological analyzers %~\cite{Sughaiyer:04}
consider the internal structure of a word and compose it from
several {\em morphemes}. 
A morpheme is a {\em stem} or an {\em affix}.
An affix is a {\em prefix, suffix,} or an {\em infix}.
Prefixes and suffixes are attached to the beginning and end of the word respectively; 
while infixes introduce changes in the stem itself. 
Morphological analyzers have been extensively used as building blocks 
in many text mining applications~\cite{Sou07}.  In Arabic, morphological analyzers such as  
Buckwalter~\cite{Buckwalter:02},
Beesley~\cite{Beesley:01},
SAMA ~\cite{Kulick:10},
ElixirFM~\cite{Otakar:07}, 
AMIRA~\cite{Diab:07,Benajiba:07},
MAGEAD~\cite{Habash:05}, 
and MADA+TOKAN~\cite{Habash:09}
exist.
They preprocess text to remove diacritics, short vowels, such as 
\RL{ha}, \RL{hu}, \RL{hi}, \RL{TaN}, \RL{huN}, \RL{hiN}.~\footnote{In this document, 
we use the default ArabTeX transliteration style ZDMG.} Diacritics help to disambiguate the
meaning of a word. For example, \noVocRL{wrd} may mean ``rose"(\RL{ward}), ``came"(\fullvocalize\RL{warada}) 
or ``and he replied"(\RL{waradda}). We think that using diacritics in the analysis can highly increase accuracy.
Current analyzers take as input whitespace-delimited tokens. However, in Arabic word boundaries can be more complex.
Arabic letters can have up to 
four different forms
corresponding to their position in a word, i.e. beginning,
middle, end of word and separate forms. 
This allows the phrase
\noTrnoVocRL{il_A\nospace almdrsT}  with no delimiter within
to be visually recognizable
as two separate words \noVocRL{il_A} (to) and \noVocRL{almdrsT} (the school).
This happens because the first word \noVocRL{il_A} ends with
\noVocRL{_A} a non-connecting letter. We propose to solve this problem, commonly known as the run-on word problem, using online tokenization.
Online tokenization also solves the issue of multi-word expressions such as \noVocRL{`abid al-la.tiyf} and 
allows them to be analyzed as one unit -- here a compound proper name.

@@

%%Similar to %the 
%%MADA+TOKAN, Sarf uses
%%the several valid morphological analyses as 
%%%a richness that should 
%%richness at a higher abstract level, and does not view them as 
%%inaccuracies to be corrected. 
%Sarf differs from MADA+TOKAN in that it provides an interface for the 
%application to interfere in judging the analyses much earlier and
%at a finer granularity level. 
%%In Sarf, there is no need for a separate tokenizer such as
%Sarf does not need a separate tokenizer such as
%TOKAN since each solution keeps a stack of positions
%that partition text into morphemes.
%%The application-specific controller can build the trace 
%%when needed.

@@

The main tasks of the thesis are listed below: \\
\begin{enumerate}[1.]
\item \textbf{Build and Enhance Sarf} \label{I:BuildSarf}
	\begin{enumerate}[\ref*{I:BuildSarf}.1.]
	\item Port Buckwalter into C++ and mysql \label{I:portBuckwalter}
	\item Design and implement the Non-deterministic Finite State Automata framework \label{I:ndfsa}
	\item Implement an application-specific controller interface \label{I:applicationController}
	\item Support for disambiguation using diacritics \label{I:diacritics}
	\item Support for multi-word expressions \label{I:multiword}
	\item Solve the issue of `run-on' words \label{I:runon}
	\item Validate Sarf \label{I:validateSarf}
	\item Benchmark against existing corpora \label{I:benchMark}
	\end{enumerate}
\item \textbf{Support Sarf} \label{I:SupportSarf}
	\begin{enumerate}[\ref*{I:SupportSarf}.1.]
	\item Optimize performance \label{I:optimizeSarf}
	\item Support for running Sarf under Windows \label{I:SarfonWindows}
	\item Refactor the code \label{I:RefactorSarf}
	\item Write a user Manual \label{I:SarfUserManual}
	\item Write a developer manual and documentation \label{I:SarfDeveloperManual}
	\end{enumerate}
\item \textbf{Build a time entity extraction application} \label{I:TimeEntityExtractor}
	\begin{enumerate}[\ref*{I:TimeEntityExtractor}.1.]
	\item Recognize time entities \label{I:TimeEntityRecognition}
	\item Normalize extracted time entities \label{I:TimeEntityNormalization}
	\item Validate with a sorting application \label{I:TimeEntityValidation}
	\end{enumerate}
\item \textbf{Automate Hadith Authentication} \label{I:Hadith}
	\begin{enumerate}[\ref*{I:Hadith}.1.]
	\item Extend the Lexicon of Buckwalter with proper names \label{I:HadithNames}
	\item Implement the hadith segmentation tool \label{I:HadithSegmentation}
	\item Write the Narrator Equality Function \label{I:HadithEquality}
	\item Build Parial Order Graphs \label{I:HadithPOR}
	\item Correct Graph mistakes that result from Ambiguous narrator names which introduce false cycles to the graph \label{I:HadithBreakCycles}
	\item Segment biography books using a graph coloring algorithm \label{I:BiographySegmentation}
	\item Extract Annotations from the biography books \label{I:BiographyAnnotationExtraction}
	\item Assign annotations to graph nodes \label{I:HadithPORAnnotation}
	\item Enable and perform consistency checks between narrator chains \label{I:HadithChecks}
	\item Perform experiments on the effects of different narrators in hadith literature \label{I:HadithExperiments}
	\item Write a user manual \label{I:HadithManual}
	\item Distribute the tool to volunteer scholars and report their findings \label{I:HadithScolars}
	\end{enumerate}
\item \textbf{Write thesis Report and defend the thesis} \label{I:WriteReport}
\end{enumerate}

@@

\begin{table}[bt]
\centering
\caption{Thesis Time Table}
\resizebox{1\columnwidth}{!}{
\begin{tabular}{|c|c||c|c||c|c||c||c|c|c|c||c|c|c|c||c|} \bottomrule[0.1em]
\textbf{Year} & \multicolumn{9}{|c|}{\textbf{2011}} & \multicolumn{6}{|c|}{\textbf{2012}} \\ \hline
\textbf{Semester} & ~ & \multicolumn{2}{|c||}{\textbf{\tiny Spring}} & \multicolumn{2}{|c||}{\textbf{\tiny Summer}} 
		& ~ & \multicolumn{4}{|c||}{\textbf{\tiny Fall}} & \multicolumn{4}{|c||}{\textbf{ \tiny Spring}} & ~ \\ \hline
\backslashbox{Task}{Month}
			& \textbf{Done} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8} & \textbf{9} & \textbf{\small 10} & 
			 \textbf{\small 11} & \textbf{\small 12} &
			 \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} &
				\begin{tabular}{c}
				\textbf{\small Future} \\
				\textbf{\small Work}
				\end{tabular}
			 \\ \toprule[0.1em] \bottomrule[0.1em]
\textbf{\ref{I:portBuckwalter}} 				& \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:ndfsa}} 							& \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ &\\ \hline
\textbf{\ref{I:applicationController}} 			& \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:diacritics}}						& \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ &\\ \hline
\textbf{\ref{I:multiword}} 						& \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:runon}} 							& \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:benchMark}} 						& \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ &\\ \hline
\textbf{\ref{I:validateSarf}} 					& ~ & ~ & ~ & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ &\\ \toprule[0.1em] \bottomrule[0.1em]
\textbf{\ref{I:optimizeSarf}} 					& \x & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ &\X & ~ &\X &\\ \hline
\textbf{\ref{I:SarfonWindows}} 					& \x & ~ &\X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:RefactorSarf}} 					& ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X &\X & ~ & \\ \hline
\textbf{\ref{I:SarfUserManual}}					& ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X & ~ & \\ \hline
\textbf{\ref{I:SarfDeveloperManual}}			& ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X & ~ & \\ \toprule[0.1em] \bottomrule[0.1em]
\textbf{\ref{I:TimeEntityRecognition}}  		& \x &\X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:TimeEntityNormalization}}		& ~ & \X &\X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline 
\textbf{\ref{I:TimeEntityValidation}}			& ~ & ~ &\X & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \toprule[0.1em] \bottomrule[0.1em]
\textbf{\ref{I:HadithNames}} 					& \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline 
\textbf{\ref{I:HadithSegmentation}} 			& \x & ~ & ~ & ~ &\X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:HadithEquality}} 				& \x & ~ & ~ & ~ &\X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:HadithPOR}} 						& \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:HadithBreakCycles}} 				& \x & ~ & ~ & ~ & ~ &\X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:BiographySegmentation}} 			& ~ & ~ & ~ & ~ & ~ &\X & \X & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:BiographyAnnotationExtraction}} 	& ~ & ~ & ~ & ~ & ~ & ~ &\X &\X & \X & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:HadithPORAnnotation}} 			& ~ & ~ & ~ & ~ & ~ & ~ & ~ &\X & \X & ~ & ~ & ~ & ~ & ~ & \\ \hline
\textbf{\ref{I:HadithChecks}} 					& ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ &\X &\X & \X & ~ & ~ & ~ & \X \\ \hline
\textbf{\ref{I:HadithExperiments}} 				& ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ &\X & \X & ~ & ~ & \X \\ \hline
\textbf{\ref{I:HadithManual}} 					& ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X & ~ & \\ \hline
\textbf{\ref{I:HadithScolars}} 					& ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~  & ~ & \X \\ \toprule[0.1em] \bottomrule[0.1em]
\textbf{\ref{I:WriteReport}} 					& ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & ~ & \X &  \\ \hline
\end{tabular}
}
\normalsize
\label{t:timetable}
\end{table}
